---
title: "Assignment_3"
author: "Patrik Molnar"
date: '2022-11-09'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Assignment 3
```{r stup}
pacman::p_load(glue,
       tidyr,
       data.table,
       moments,
       tidybayes,
       tibble,
       cowplot,
       viridis,
       brms,
       rstan,
       cmdstanr,
       magrittr,
       gridExtra,
       grid,
       lattice,
       ggplot2,
       ggridges,
       ellipse,
       Rmisc, 
       dplyr, 
       "rmarkdown", 
       knitr,
       yardstick,
       purr,
       splitstackshape,
       tidymodels)
pacman::p_load(tidyverse)
```
##Part 1 - Simulating data
 
Use meta analysis reported in Parola et al (2020) to create informed simulated data 
- 100 pairs of schizophrenia and controls, each participant producing 10 repeated measures (10 trials with their speech recorded), for each recording produce 10 acoustic measures (6 from meta analysis and 4 with random noise)

- Do the same for a baseline data set including only 10 noise variables 

### seeting up variables
```{r}
###Data simulation 
n <- 100
trials <- 10

#Effect sizes definition = Informed effect mean and Skeptic effect mean
IEM <- c(-0.5,-1.26,-.74,1.89,0.25,1.3,0,0,0,0)
SEM <- rep(0,10)

#Defining individual vairability from populationd accross trials measurement error
ISD <- 1
TSD <- 0.5
E <- 0.2

```
### Simulating the true effect size for each varibale for all pairs of participants
```{r}
for (i in seq(10)){
  temp_informed <- tibble(
    ID=seq(n),
    TrueEffect = rnorm(n, IEM[i], ISD),
    Variable = paste0("V",i))
  temp_skeptic <- tibble(
    ID=seq(n),
    TrueEffect = rnorm(n, SEM[i], ISD),
    Variable = paste0("V",i))
  
  if(i==1){
    d_informed_true <- temp_informed
    d_skeptic_true <- temp_skeptic
  } else {
    d_informed_true <-  rbind(d_informed_true, temp_informed)
    d_skeptic_true <- rbind(d_skeptic_true, temp_skeptic)
  }
}
```
### Creating one row per trial 
```{r}
d_trial <- tibble(expand_grid(ID=seq(n), Trial = seq(trials), Group = c("Schizophrenia", "Control")))

d_informed <- merge(d_informed_true, d_trial)
d_skeptic <- merge(d_skeptic_true, d_trial)

for ( i in seq(nrow(d_informed))){
  d_informed$measurement[i] <- ifelse(d_informed$Group[i]=="Schizophrenia",
                                      rnorm(1, rnorm(1, d_informed$TrueEffect[i]/2, TSD), E),
                                      rnorm(1, rnorm(1, (-d_informed$TrueEffect[i])/2, TSD), E))
  
  d_skeptic$measurement[i] <- ifelse(d_skeptic$Group[i]=="Schizophrenia",
                                      rnorm(1, rnorm(1, d_skeptic$TrueEffect[i]/2, TSD), E),
                                      rnorm(1, rnorm(1, (-d_skeptic$TrueEffect[i])/2, TSD), E))
}


```
### Transforming the dataframe to a wide format based on the variable
```{r}
d_informed_wide <- d_informed %>% 
  mutate(TrueEffect=NULL) %>% 
  pivot_wider(names_from = Variable,
              values_from = measurement)
d_skeptic_wide <- d_skeptic %>% 
  mutate(TrueEffect=NULL) %>% 
  pivot_wider(names_from = Variable,
              values_from = measurement)

Schizo_ID <- d_informed_wide %>% 
  filter(Group == 'Schizophrenia')
 
 control_ID <- d_informed_wide %>% 
  filter(Group == 'Control')
 
control_ID[, 1] <- control_ID[,1] + 100

 d_informed_wide <- rbind(control_ID,Schizo_ID)


 Schizo_ID_skeptic <- d_skeptic_wide %>% 
  filter(Group == 'Schizophrenia')
 
 control_ID_skeptic <- d_skeptic_wide %>% 
  filter(Group == 'Control')
 
 control_ID_skeptic[,1] <- control_ID_skeptic[,1] + 100
 
 d_skeptic_wide <- rbind(control_ID_skeptic,Schizo_ID_skeptic)
```


### Visualizing the simulated informed data
```{r}
plot1 <- d_informed_wide %>% 
  ggplot(aes(x = V1, color = Group))+
  geom_density()

plot2 <- d_informed_wide %>% 
  ggplot(aes(x = V2, color = Group))+
  geom_density()

plot3 <- d_informed_wide %>% 
  ggplot(aes(x = V3, color = Group))+
  geom_density()

plot4 <- d_informed_wide %>% 
  ggplot(aes(x = V4, color = Group))+
  geom_density()

plot5 <- d_informed_wide %>% 
  ggplot(aes(x = V5, color = Group))+
  geom_density()

plot6 <- d_informed_wide %>% 
  ggplot(aes(x = V6, color = Group))+
  geom_density()

plot7 <- d_informed_wide %>% 
  ggplot(aes(x = V7, color = Group))+
  geom_density()

plot8 <- d_informed_wide %>% 
  ggplot(aes(x = V8, color = Group))+
  geom_density()

plot9 <- d_informed_wide %>% 
  ggplot(aes(x = V9, color = Group))+
  geom_density()

plot10 <- d_informed_wide %>% 
  ggplot(aes(x = V10, color = Group))+
  geom_density()

cowplot::plot_grid(plot1, plot2, plot3, plot4, plot5, plot6, plot7, plot8, plot9, plot10, 
          labels = c("v1", "v2", "v3",'v4','v5','v6','v7','v8','v9','v10'),
          ncol = 5, nrow = 2)


```

### Visualizing the simulated data for skeptical data frame
```{r}
plot1_s <- d_skeptic_wide %>% 
  ggplot(aes(x = V1, color = Group))+
  geom_density()

plot2_s <- d_skeptic_wide %>% 
  ggplot(aes(x = V2, color = Group))+
  geom_density()

plot3_s <- d_skeptic_wide %>% 
  ggplot(aes(x = V3, color = Group))+
  geom_density()

plot4_s <- d_skeptic_wide %>% 
  ggplot(aes(x = V4, color = Group))+
  geom_density()

plot5_s <- d_skeptic_wide %>% 
  ggplot(aes(x = V5, color = Group))+
  geom_density()

plot6_s <- d_skeptic_wide %>% 
  ggplot(aes(x = V6, color = Group))+
  geom_density()

plot7_s <- d_skeptic_wide %>% 
  ggplot(aes(x = V7, color = Group))+
  geom_density()

plot8_s <- d_skeptic_wide %>% 
  ggplot(aes(x = V8, color = Group))+
  geom_density()

plot9_s <- d_skeptic_wide %>% 
  ggplot(aes(x = V9, color = Group))+
  geom_density()

plot10_s <- d_skeptic_wide %>% 
  ggplot(aes(x = V10, color = Group))+
  geom_density()

cowplot::plot_grid(plot1_s, plot2_s, plot3_s, plot4_s, plot5_s, plot6_s, plot7_s, plot8_s, plot9_s, plot10_s, 
          labels = c("v1", "v2", "v3",'v4','v5','v6','v7','v8','v9','v10'),
          ncol = 5, nrow = 2)


```

##Part 2 - Machine larning pipeline on simulated data 
Build a machine leaning pipeline (separately on the 2 datasets)
- create a data budget (e.g., balanced training and test sets)
pree-process the data (e.g., scaling the features)
- fit and assess a classification algorithm on the training data (e.g., bayesian multilevel logistic regression)
- assess performance on the test set
- discuss whether performance and feature importance is as expected

### Informed dataset
#### Data budget and pree-processing (scaling)
##### splitting the dataset into training and test
```{r}
##Machine Learning pipeline

d_informed_wide <- d_informed_wide %>% 
  mutate(pair_ID=ID) %>% 
  mutate(pair_ID= ifelse(ID>100, ID-100, ID))

d_skeptic_wide <- d_skeptic_wide %>% 
  mutate(pair_ID=ID) %>% 
  mutate(pair_ID= ifelse(ID>100, ID-100, ID))

sample_inf <- sample(seq(100), 80)
train_informed <- subset(d_informed_wide, pair_ID %in% sample_inf)
test_informed <- subset(d_informed_wide, !pair_ID %in% sample_inf)

sample_inf <- sample(seq(100), 80)
train_skeptic <- subset(d_skeptic_wide, pair_ID %in% sample_skep)
test_skeptic <- subset(d_skeptic_wide, !pair_ID %in% sample_skep)

train_informed$ID <- as.factor(train_informed$ID)
train_skeptic$ID <- as.factor(train_skeptic$ID)

test_informed$ID <- as.factor(test_informed$ID)
test_skeptic$ID <- as.factor(test_skeptic$ID)

train_informed$pair_ID <- as.factor(train_informed$pair_ID)
train_skeptic$pair_ID <- as.factor(train_skeptic$pair_ID)

test_informed$pair_ID <- as.factor(test_informed$pair_ID)
test_skeptic$pair_ID <- as.factor(test_skeptic$pair_ID)

train_informed$Trial <- as.factor(train_informed$Trial)
train_skeptic$Trial <- as.factor(train_skeptic$Trial)

test_informed$Trial <- as.factor(test_informed$Trial)
test_skeptic$Trial <- as.factor(test_skeptic$Trial)

##Splitting the data using stratified, which also take into consideration stratification of the f.e. gender


#Scaling
rec_informed <- train_informed %>% 
  recipe(Group~.) %>% 
  update_role(ID, new_role = 'ID') %>% 
  step_scale(all_numeric()) %>% 
  step_center(all_numeric()) %>% 
  prep(training=train_informed, retain=TRUE)

rec_skeptic <- train_skeptic %>% 
  recipe(Group~.) %>% 
  update_role(ID, new_role = 'ID') %>% 
  step_scale(all_numeric()) %>% 
  step_center(all_numeric()) %>% 
  prep(training=train_informed, retain=TRUE)


train_informed_s <- juice(rec_informed)
test_informed_s <- bake(rec_informed, new_data = test_informed)

train_skeptic_s <- juice(rec_skeptic)
test_skeptic_s <- bake(rec_skeptic, new_data = test_skeptic)

#Ask ricardo about the ID scaling
```
#### Visual inspection of the data
```{r}
#Plotting
plot_i2 <- ggplot(train_informed, aes(V2, Group, colour=Group))+
  geom_point()+
  geom_smooth(method = "glm", se=FALSE)+
  theme_bw()+
  ggtitle("informed")

plot_s2 <- ggplot(train_skeptic, aes(V2, Group, colour=Group))+
  geom_point()+
  geom_smooth(method = "glm", se=FALSE)+
  theme_bw()+
  ggtitle("sceptical")

plot_i3 <- ggplot(train_informed, aes(V3, Group, colour=Group))+
  geom_point()+
  geom_smooth(method = "glm", se=FALSE)+
  theme_bw()+
  ggtitle("informed")

plot_s3 <- ggplot(train_skeptic, aes(V3, Group, colour=Group))+
  geom_point()+
  geom_smooth(method = "glm", se=FALSE)+
  theme_bw()+
  ggtitle("sceptical")

plot_i4 <- ggplot(train_informed, aes(V4, Group, colour=Group))+
  geom_point()+
  geom_smooth(method = "glm", se=FALSE)+
  theme_bw()+
  ggtitle("informed")

plot_s4 <- ggplot(train_skeptic, aes(V4, Group, colour=Group))+
  geom_point()+
  geom_smooth(method = "glm", se=FALSE)+
  theme_bw()+
  ggtitle("sceptical")

plot_i6 <- ggplot(train_informed, aes(V6, Group, colour=Group))+
  geom_point()+
  geom_smooth(method = "glm", se=FALSE)+
  theme_bw()+
  ggtitle("informed")

plot_s6 <- ggplot(train_skeptic, aes(V6, Group, colour=Group))+
  geom_point()+
  geom_smooth(method = "glm", se=FALSE)+
  theme_bw()+
  ggtitle("sceptical")

plot_grid(plot_i2, plot_s2, plot_i3, plot_s3, plot_i4, plot_s4, plot_i6, plot_s6)

#Make the distribution 
```
####fit and asses a classification algorithm on training data (Baysian)

```{r}
##Setting up the model
PR_f0 <-bf(Group~1+V1+V2+V3+V4+V5+V6+V7+V8+V9+V10)

PR_f1 <-bf(Group~1+V1+V2+V3+V4+V5+V6+V7+V8+V9+V10+(1+Trial|ID))

get_prior(PR_f0, train_informed_s, family = bernoulli)

get_prior(PR_f1, train_informed_s, family = bernoulli)
```
```{r}
##setting the priors
PR_p0 <- c(
  prior(normal(0, 1), class=Intercept),
  prior(normal(0, 0.3), class=b)
)

PR_p1 <- c(
  prior(normal(0, 1), class=Intercept),
  prior(normal(0, 0.3), class=sd),
  prior(normal(0, 0.3), class=b),
  prior(lkj(1), class=cor)
)
```

##### Fitting the first model on both the skeptical and informed data

```{r}
##Model fit on priors
pr_m0_inf <- brm(
  PR_f0, 
  data = train_informed_s, 
  prior = PR_p0,
  family = bernoulli,
  refresh=0,
  sample_prior = 'only',
  iter=6000,
  warmup = 2500,
  backend = "cmdstanr",
  threads = threading(2),
  chains = 4,
  cores = 4,
  control = list(
    adapt_delta = 0.9,
    max_treedepth = 20)
)


pp_check(pr_m0_inf)

pr_m0_skep <- brm(
  PR_f0, 
  data = train_skeptic_s, 
  prior = PR_p0,
  family = bernoulli,
  refresh=0,
  sample_prior = 'only',
  iter=6000,
  warmup = 2500,
  backend = "cmdstanr",
  threads = threading(2),
  chains = 4,
  cores = 4,
  control = list(
    adapt_delta = 0.9,
    max_treedepth = 20)
)


pp_check(pr_m0_skep)
```

```{r}
##Model fit on informed data
pr_m0_fit_inf <- brm(
  PR_f0, 
  data = train_informed_s, 
  prior = PR_p0,
  family = bernoulli,
  refresh=0,
  sample_prior = TRUE,
  iter=6000,
  warmup = 2500,
  backend = "cmdstanr",
  threads = threading(2),
  chains = 4,
  cores = 4,
  control = list(
    adapt_delta = 0.9,
    max_treedepth = 20)
)

pp_check(pr_m0_fit_inf)

plot(pr_m0_fit_inf)

summary(pr_m0_fit_inf)

```

```{r}
##Model fit on skeptical data
pr_m0_fit_skep <- brm(
  PR_f0, 
  data = train_skeptic_s, 
  prior = PR_p0,
  family = bernoulli,
  refresh=0,
  sample_prior = TRUE,
  iter=6000,
  warmup = 2500,
  backend = "cmdstanr",
  threads = threading(2),
  chains = 4,
  cores = 4,
  control = list(
    adapt_delta = 0.9,
    max_treedepth = 20)
)

pp_check(pr_m0_fit_skep)

plot(pr_m0_fit_skep)

summary(pr_m0_fit_skep)

```
##### Fitting the second model on skeptical and informed data
```{r}
##Second model fit on priors
pr_m1_inf <- brm(
  PR_f1, 
  data = train_informed_s, 
  prior = PR_p1,
  family = bernoulli,
  refresh=0,
  sample_prior = 'only',
  iter=6000,
  warmup = 2500,
  backend = "cmdstanr",
  threads = threading(2),
  chains = 4,
  cores = 4,
  control = list(
    adapt_delta = 0.9,
    max_treedepth = 20)
)


pp_check(pr_m1_inf)

pr_m1_skep <- brm(
  PR_f1, 
  data = train_skeptic_s, 
  prior = PR_p1,
  family = bernoulli,
  refresh=0,
  sample_prior = 'only',
  iter=6000,
  warmup = 2500,
  backend = "cmdstanr",
  threads = threading(2),
  chains = 4,
  cores = 4,
  control = list(
    adapt_delta = 0.9,
    max_treedepth = 20)
)


pp_check(pr_m1_skep)
```
```{r}
##second model fit on informed data
pr_m1_fit_inf <- brm(
  PR_f1, 
  data = train_informed_s, 
  prior = PR_p1,
  family = bernoulli,
  refresh=0,
  sample_prior = TRUE,
  iter=6000,
  warmup = 2500,
  backend = "cmdstanr",
  threads = threading(2),
  chains = 4,
  cores = 4,
  control = list(
    adapt_delta = 0.9,
    max_treedepth = 20)
)

pp_check(pr_m1_fit_inf)

plot(pr_m1_fit_inf)

summary(pr_m1_fit_inf)

```
```{r}
##Model fit on skeptical data
pr_m1_fit_skep <- brm(
  PR_f1, 
  data = train_skeptic_s, 
  prior = PR_p1,
  family = bernoulli,
  refresh=0,
  sample_prior = TRUE,
  iter=6000,
  warmup = 2500,
  backend = "cmdstanr",
  threads = threading(2),
  chains = 4,
  cores = 4,
  control = list(
    adapt_delta = 0.9,
    max_treedepth = 20)
)

pp_check(pr_m1_fit_skep)

plot(pr_m1_fit_skep)

summary(pr_m1_fit_skep)

```


#### Asses performance on test data
```{r}
train_informed_s$PredictionsPerc0 <- predict(pr_m0_fit_inf)[, 1]
train_informed_s$Prediction0[train_informed_s$PredictionsPerc0 > 0.5] <- "Schizophrenia" 
train_informed_s$Prediction0[train_informed_s$PredictionsPerc0 <= 0.5] <- "Control" 


train_informed_s$PredictionsPerc1 <- predict(pr_m1_fit_inf)[, 1]
train_informed_s$Prediction1[train_informed_s$PredictionsPerc1 > 0.5] <- "Schizophrenia" 
train_informed_s$Prediction1[train_informed_s$PredictionsPerc1 <= 0.5] <- "Control" 


train_skeptic_s$PredictionsPerc0 <- predict(pr_m0_fit_skep)[, 1]
train_skeptic_s$Prediction0[train_skeptic_s$PredictionsPerc0 > 0.5] <- "Schizophrenia" 
train_skeptic_s$Prediction0[train_skeptic_s$PredictionsPerc0 <= 0.5] <- "Control" 

train_skeptic_s$PredictionsPerc1 <- predict(pr_m1_fit_skep)[, 1]
train_skeptic_s$Prediction1[train_skeptic_s$PredictionsPerc1 > 0.5] <- "Schizophrenia" 
train_skeptic_s$Prediction1[train_skeptic_s$PredictionsPerc1 <= 0.5] <- "Control" 

train_informed_s <- train_informed_s %>% 
  mutate(
    Group = as.factor(Group),
    Prediction0 = as.factor(Prediction0),
    Prediction1 = as.factor(Prediction1)
  )

train_skeptic_s <- train_skeptic_s %>% 
  mutate(
    Group = as.factor(Group),
    Prediction0 = as.factor(Prediction0),
    Prediction1 = as.factor(Prediction1)
  )


test_informed_s$PredictionsPerc0 <- predict(pr_m0_fit_inf, newdata = test_informed_s, allow_new_levels = T)[, 1]
test_informed_s$Prediction0[test_informed_s$PredictionsPerc0 > 0.5] <- "Schizophrenia" 
test_informed_s$Prediction0[test_informed_s$PredictionsPerc0 <= 0.5] <- "Control" 


test_informed_s$PredictionsPerc1 <- predict(pr_m1_fit_inf, newdata = test_informed_s, allow_new_levels = T)[, 1]
test_informed_s$Prediction1[test_informed_s$PredictionsPerc1 > 0.5] <- "Schizophrenia" 
test_informed_s$Prediction1[test_informed_s$PredictionsPerc1 <= 0.5] <- "Control" 


test_skeptic_s$PredictionsPerc0 <- predict(pr_m0_fit_skep, newdata = test_informed_s, allow_new_levels = T)[, 1]
test_skeptic_s$Prediction0[test_skeptic_s$PredictionsPerc0 > 0.5] <- "Schizophrenia" 
test_skeptic_s$Prediction0[test_skeptic_s$PredictionsPerc0 <= 0.5] <- "Control" 

test_skeptic_s$PredictionsPerc1 <- predict(pr_m1_fit_skep, newdata = test_informed_s, allow_new_levels = T)[, 1]
test_skeptic_s$Prediction1[test_skeptic_s$PredictionsPerc1 > 0.5] <- "Schizophrenia" 
test_skeptic_s$Prediction1[test_skeptic_s$PredictionsPerc1 <= 0.5] <- "Control"

test_informed_s <- test_informed_s %>% 
  mutate(
    Group = as.factor(Group),
    Prediction0 = as.factor(Prediction0),
    Prediction1 = as.factor(Prediction1)
  )

test_skeptic_s <- test_skeptic_s %>% 
  mutate(
    Group = as.factor(Group),
    Prediction0 = as.factor(Prediction0),
    Prediction1 = as.factor(Prediction1)
  )
```
```{r}
conf_mat(train_informed_s, 
         truth = Group, 
         estimate = Prediction0, 
         dnn = c('prediction','truth'))

conf_mat(test_informed_s, 
         truth = Group, 
         estimate = Prediction0, 
         dnn = c('prediction','truth'))

conf_mat(train_informed_s, 
         truth = Group, 
         estimate = Prediction1, 
         dnn = c('prediction','truth'))

conf_mat(test_informed_s, 
         truth = Group, 
         estimate = Prediction1, 
         dnn = c('prediction','truth'))

conf_mat(train_skeptic_s, 
         truth = Group, 
         estimate = Prediction0, 
         dnn = c('prediction','truth'))

conf_mat(test_skeptic_s, 
         truth = Group, 
         estimate = Prediction0, 
         dnn = c('prediction','truth'))

conf_mat(train_skeptic_s, 
         truth = Group, 
         estimate = Prediction1, 
         dnn = c('prediction','truth'))

conf_mat(test_skeptic_s, 
         truth = Group, 
         estimate = Prediction1, 
         dnn = c('prediction','truth'))
```


#### Discuss whether performance and feature importance is as expected


##Part 3 - Applying the machine learning pipeline to empirical data 

```{r}
real_data <- read_csv("assignment_3_data.csv")
```


Apply your machine learning pipeline to the empirical data 

Warning: in simulated data you only ahve 10 features, now you have many more 
- Consider the impact a higher number of features will have on your ML inference and decide if you want to cut down the number of features before running the pipeline (alternatibvely expand the pipeline to add feature selection)